---
title: "Introduction to Research Methods"
author: "Stijn Masschelein"
date: "`r format(Sys.Date(), '%B, %Y')`"
output:
  html_document: 
    df_print: kable
    number_sections: yes
    theme: readable
    css: styles.css
    toc: true
  ioslides_presentation: default
---

<!--
Things I did to make this work
- See yaml: The html document are the notes, the slides are well the slides
- Made a snippet in Rstudio (`no`) for the notes div; (`cs`) for the content
  of slides.
- Made an additional style for the notes and the slides.
- This YAML section need to come first
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Research Methods in Finance and Accounting

## Introduction

<div class="content_slide">
- Preparation to do empirical research (in accounting and finance)
- From theory to data
- From data to statistical tests
</div>

<div class="notes">
The emphasis on the units will be on the practical and statistical issues of
the data analysis part of a thesis. The influence will not be on the accounting
and finance part of your thesis. The goal is to be relevant to everyone in the
unit.
</div>

## Four Modules

<div class="content_slide">
1. The research process and basic data skills by me
2. Timeseries and databases by Prof. Dirk Bauer
3. Panel data and identification by Prof. Jing Yang
4. Restricted dependent variable and survey data by Prof. Paul Gerrans
</div>

<div class="notes">
All modules work as stand-alone 3-week units and aim to cover a wide
range of topics. Not all of them will in the end be relevant for everyone. 
However, it is probably a good idea to get to know the different statistical
methods, their advantages, and disadvantages. Bringing a new methodology to 
an old topic can be a valuable contribution. Some problems have already been
solved in other research streams.
</div>

## Assessment

<div class="content_slide">
- No Exam
- Four assignments (%)
- Proposal and presentation (%) 
</div>

<div class="notes">
We want you to (1) do some data analysis and (2) be well prepared to undertake
(the data analaysis part of) a research project. So we are going to evaluate 
you by letting you (1) analyse data and (2) prepare your honours thesis. 
</div>

# The first three weeks: CEO compensation
## Warning: I am not a specialist

```{r out.width = "100%"}
knitr::include_graphics(
  "http://i0.kym-cdn.com/entries/icons/original/000/008/342/ihave.jpg")
```

<div class="notes">
I am not a specialist in the topic nor in this type of data analysis. CEO
compensation is something that people in finance, accounting, economics, and
outside of academia are interested in. The topic is probably the one with the
most commonality. I am comforable with these type of economic theories and I am
going to stress the role of theory in data analysis a lot. Some of you will have
a topic that is at first sight less theory driven or rely more strongly on very
specific knowledge about your setting.
</div>

## Topic

<div class="content_slide">
Completely based on ["Executive Compensation: A Modern Primer"](https://dx.doi.org/10.1257/jel.20161153) by Alex Edmans and
Xavier Gabraix in *Journal of Economic Literature*.

- The level of CEO compensation
- CEO incentives
</div>

<div class="notes">
I am going to focus on two topics.

1. How high can we expect the total compensation of a CEO to be (compared to 
   other CEOs) based on some simple economic assumptions. Too high CEO 
   compensation is sometimes seen as a signal of bad corporate goverance. 
   To measure what 'too high' means, we first need to establish a baseline of
   normal levels of compensation.
2. How should CEOs be incentivised: equity or options? How schould we measure
   whether CEOs have appropriate incentives: \$ for \$ increases, % for % 
   increases? 
</div>

## Title




